---
title: "Reproducible Research: Peer Assessment 1"
author: "elm"
output: html_document
keep_md: true
---

*Setup so code is readable, install packages*  

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)  

library(ggplot2)
library(dplyr)
```

## Loading and preprocessing the data  
*Load data, aggregate by dates*  

```{r eval=TRUE}
activity <- read.csv("activity.csv")  

daysteps <- aggregate(activity$steps, by=list(date=activity$date), FUN=sum)
```

## What is mean total number of steps taken per day?  
*Calculate mean and median steps per day*  

```{r eval=TRUE}
meansteps <- mean(daysteps$x, na.rm=TRUE)
meansteps  

mediansteps <- median(daysteps$x, na.rm=TRUE)
mediansteps
```

*Plot histogram of steps by day*  

```{r eval=TRUE}
ggplot(daysteps, aes(x=date, y=x)) +
  geom_histogram(stat="identity", width=1) + 
  theme(legend.position = "none") +
  labs(y="Step count", x="Date")
```

## What is the average daily activity pattern?  
*Average intervals across days, find interval with highest number of steps*  

```{r eval=TRUE}
intervalsteps <- aggregate(activity$steps, by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)  

highest <- intervalsteps[which.max(intervalsteps$x),]
highest
```

*Plot histogram of steps by interval*  

```{r eval=TRUE}
ggplot(intervalsteps, aes(x=interval, y=x)) +
  geom_histogram(stat="identity", width=5) + 
  theme(legend.position = "none") +
  labs(y="Step count", x="Interval")
```

## Imputing missing values  
*Replace NAs in dataset*  

```{r eval=TRUE}
sum(!complete.cases(activity))
meaninterval <- mean(activity$steps, na.rm=TRUE)
activity$steps[is.na(activity$steps)] <- meaninterval
```

*Redo histogram, mean and median of steps per day*  

```{r eval=TRUE}
daysteps2 <- aggregate(activity$steps, by=list(date=activity$date), FUN=sum)
ggplot(daysteps2, aes(x=date, y=x)) +
  geom_histogram(stat="identity", width=1) + 
  theme(legend.position = "none") +
  labs(y="Step count", x="Date")  

meansteps2 <- mean(daysteps2$x, na.rm=TRUE)
meansteps2  

mediansteps2 <- median(daysteps2$x, na.rm=TRUE)
mediansteps2
```

*Histogram is similar with fewer gaps. Mean and median are the same.*  

## Are there differences in activity patterns between weekdays and weekends?  
*Add column to activity dataframe with weekday/weekend*

```{r eval=TRUE}
activity <- mutate(activity, date=as.POSIXlt(date, format="%Y-%m-%d"))
activity$weekday = weekdays(activity$date)  

activity$weekday[activity$weekday=="Monday"|activity$weekday=="Tuesday"|activity$weekday=="Wednesday"|activity$weekday=="Thursday"|activity$weekday=="Friday"] <- "Weekday"
activity$weekday[activity$weekday=="Saturday"|activity$weekday=="Sunday"] <- "Weekend"  

weekdays <- activity[activity$weekday=="Weekday",]
weekends <- activity[activity$weekday=="Weekend",]  

weekdays2 <- aggregate(weekdays$steps, list(weekdays$interval), FUN=mean, na.rm=TRUE)
weekdays2$weekday <- "Weekday"
weekends2 <- aggregate(weekends$steps, list(weekends$interval), FUN=mean, na.rm=TRUE)
weekends2$weekday <- "Weekend"
activity2 <- rbind(weekdays2, weekends2)  

p <- ggplot(activity2, aes(x=Group.1, y=x)) + 
  geom_line() + 
  facet_wrap(~weekday, nrow=2) +
  xlab("Time interval") + 
  ylab("Steps")
p

```
